---
title: "bibr — R API Client Demo"
format:
  html:
    df-print: kable
engine: knitr
page-layout: full
toc: true
---

This notebook demonstrates how to call the bibr REST API from R.

**Prerequisites:**

- bibr API server running (`uv run bibr serve` or via Docker)
- R packages: `httr2`, `jsonlite`

## Configuration

```{r}
library(httr2)
library(jsonlite)

`%||%` <- function(x, fallback) if (is.null(x)) fallback else x

# API base URL — change if your server is running elsewhere
BASE_URL <- Sys.getenv("BIBR_API_URL", "http://localhost:8000")
API_KEY <- Sys.getenv("BIBR_API_KEY")  # Match the server's AUTH_API_KEY
FILE_PATH <- Sys.getenv("PAPER_PATH", "paper.pdf")

# Helper to add auth if credentials are set
add_auth <- function(req) {
  if (nzchar(API_KEY)) req_auth_bearer_token(req, API_KEY) else req
}
```

## 1. Health check

Verify the server is running and ready.

```{r}
health <- request(BASE_URL) |>
  req_url_path_append("health") |>
  add_auth() |>
  req_perform() |>
  resp_body_json()

cat("Health:", health$status, "\n")

ready <- request(BASE_URL) |>
  req_url_path_append("ready") |>
  add_auth() |>
  req_perform() |>
  resp_body_json()

cat("Ready: ", ready$status, "\n")
cat("Checks:", paste(names(ready$checks), ready$checks, sep = "=", collapse = ", "), "\n")
```

## 2. Extract metadata (JSON)

The `/papers/extract` endpoint returns structured metadata as JSON.

```{r}
if (!file.exists(FILE_PATH)) {
  stop("File not found: ", FILE_PATH)
}

cat("Sending", basename(FILE_PATH), "to", BASE_URL, "...\n")

req <- request(BASE_URL) |>
  req_url_path_append("papers", "extract") |>
  add_auth() |>
  req_body_multipart(
    file = curl::form_file(FILE_PATH)
  ) |>
  req_timeout(300)

resp <- req |> req_perform()
data <- resp |> resp_body_json()

cat("Status:", resp_status(resp), "\n")
```

### Paper metadata

```{r}
meta <- data$metadata
cat("Title:      ", meta$title, "\n")
cat("DOI:        ", meta$doi, "\n")
cat("Keywords:   ", paste(meta$keywords, collapse = ", "), "\n")
cat("Paper type: ", meta$paper_type %||% "(unknown)", "\n")
cat("OECD L1:    ", meta$oecd_l1 %||% "(unclassified)", "\n")
cat("OECD L2:    ", meta$oecd_l2 %||% "(unclassified)", "\n")
cat("Authors:    ", length(data$author), "\n")
cat("References: ", length(data$bib), "\n")
cat("Sections:   ", length(data$section), "\n")
cat("Xrefs:      ", length(data$xref), "\n")
```

### Authors

```{r}
authors <- do.call(rbind, lapply(data$author, function(a) {
  data.frame(
    given = a$given %||% "",
    family = a$family %||% "",
    affiliation = a$affiliation %||% "",
    email = a$email %||% NA_character_,
    orcid = a$orcid %||% NA_character_,
    corresponding = a$corresponding %||% FALSE,
    stringsAsFactors = FALSE
  )
}))
authors
```

### Sections (IMRaD classification)

```{r}
# How each section's type was decided lives under extraction$diagnostics.
scores <- data$extraction$diagnostics$section_classification %||% list()
score_of <- function(id) {
  for (row in scores) if (identical(row$section_id, id)) return(row$score %||% NA_real_)
  NA_real_
}
sections <- do.call(rbind, lapply(data$section, function(s) {
  data.frame(
    section_id = s$section_id,
    header = s$header %||% "",
    section_type = s$section_type %||% "unknown",
    score = round(score_of(s$section_id), 2),
    stringsAsFactors = FALSE
  )
}))
sections
```

### References (first 10)

```{r}
refs <- do.call(rbind, lapply(data$bib, function(r) {
  data.frame(
    bib_id = r$bib_id,
    authors = r$authors %||% "",
    year = r$year %||% NA,
    title = r$title %||% "",
    stringsAsFactors = FALSE
  )
}))
head(refs, 10)
```

## 3. Save extraction result to disk

```{r}
output_path <- tempfile(fileext = ".json")
write_json(data, output_path, auto_unbox = TRUE, pretty = TRUE)
cat("Saved JSON to:", output_path, "\n")
```

## 4. Async batch processing via the platform

For processing many papers concurrently, see the [Scienceverse platform integration](../examples/platform/r_platform_demo.qmd).
That separate service has its own `/jobs` API and credentials. bibr also has
[its own async REST endpoints](../docs/reference/rest-api.md).
