---
title: "bibr — R Library Demo"
format:
  html:
    df-print: kable
engine: knitr
page-layout: full
toc: true
---

This notebook demonstrates how to read and work with bibr's v10 JSON exports
in R. bibr outputs structured JSON — this is the integration format for
downstream R tools like [Metacheck](https://github.com/scienceverse/metacheck).

**Prerequisites:**

- A JSON export from bibr (a `.json` file)
- R packages: `jsonlite`, `dplyr`

To generate an export, run bibr from the command line:

```bash
uv run bibr chew paper.pdf -o result.json
```

…or use the Python library or REST API (see companion notebooks).

## Setup

```{r}
library(jsonlite)
library(dplyr)

`%||%` <- function(x, fallback) if (is.null(x)) fallback else x
```

## 1. Generate an export with bibr (via reticulate)

If you have bibr installed in a Python environment, you can call it directly
from R using reticulate. Otherwise, skip to Section 2 and provide your own
`.json` file.

```{r}
library(reticulate)

# Use the project's virtual environment (created by `uv sync`)
use_virtualenv("../.venv", required = TRUE)

bibr <- import("bibr")
FILE_PATH <- Sys.getenv("PAPER_PATH", "paper.pdf")
if (!file.exists(FILE_PATH)) stop("File not found: ", FILE_PATH)

# chew_file manages its event loop and releases pipeline resources.
extraction <- bibr$chew_file(FILE_PATH)
data <- extraction$data

cat("Title:", data$info$title, "\n")
cat("DOI:  ", data$info$doi, "\n")

# Persist the JSON for the rest of this notebook
export_path <- tempfile(fileext = ".json")
write_json(data, export_path, auto_unbox = TRUE, pretty = TRUE)
cat("Saved JSON export to:", export_path, "\n")
```

## 2. Read the JSON file

```{r}
# Path to the JSON file — either from Section 1 or your own file
JSON_PATH <- if (exists("export_path")) export_path else Sys.getenv("BIBR_JSON_PATH", "result.json")
```

### Helper function

```{r}
# Shared helper preserves metadata and converts record arrays to data.frames.
source("../scripts/read_json_response.R")
read_bibr <- read_bibr_json
```

### Read the file

```{r}
result <- read_bibr(JSON_PATH)
cat("bibr version:", result$info$bibr_version, "\n")
cat("Tables:      ", paste(names(result), collapse = ", "), "\n")
```

## 3. Explore the data

### Paper info

```{r}
result$info
```

### Sections

```{r}
result$section
```

### Authors

```{r}
result$author
```

### Text (sentences)

```{r}
cat("Total sentences:", nrow(result$text), "\n\n")
head(result$text, 10)
```

### Join sentences with section names

```{r}
text_with_sections <- result$text |>
  left_join(
    result$section |> select(section_id, header, section_type),
    by = "section_id"
  )

head(text_with_sections, 10)
```

### Filter by section type

```{r}
# Get all sentences from the Methods section(s)
# Note: CanonicalSection.METHODS = "method" (singular)
methods <- text_with_sections |>
  filter(section_type == "method")

cat("Sentences in Methods:", nrow(methods), "\n\n")
head(methods$text, 5)
```

### References

```{r}
cat("Total references:", nrow(result$bib), "\n\n")
result$bib |> select(bib_id, authors, year, title) |> head(10)
```

### Reference matches (Crossref)

```{r}
if (nrow(result$bib_match) > 0) {
  result$bib_match |>
    select(bib_id, service, score, title, doi) |>
    head(10)
} else {
  cat("No reference matches (enrichment may be disabled)\n")
}
```

### URLs

```{r}
result$url
```

### Tables

```{r}
cat("Total tables:", nrow(result$table), "\n\n")
if (nrow(result$table) > 0) {
  result$table
}
```

### Figures

```{r}
cat("Total figures:", nrow(result$figure), "\n\n")
if (nrow(result$figure) > 0) {
  result$figure
}
```

### Cross-references

```{r}
cat("Total cross-references:", nrow(result$xref), "\n\n")
if (nrow(result$xref) > 0) {
  head(result$xref, 10)
}
```

## 4. Quick summary

```{r}
cat("Paper:          ", result$info$title %||% "(no title)", "\n")
cat("DOI:            ", result$info$doi %||% "(no DOI)", "\n")
cat("Sentences:      ", nrow(result$text), "\n")
cat("Sections:       ", nrow(result$section), "\n")
cat("Authors:        ", nrow(result$author), "\n")
cat("References:     ", nrow(result$bib), "\n")
cat("URLs:           ", nrow(result$url), "\n")
cat("Tables:         ", nrow(result$table), "\n")
cat("Figures:        ", nrow(result$figure), "\n")
cat("Cross-refs:     ", nrow(result$xref), "\n")
```
