---
title: "Scienceverse Platform — R API Client Demo"
format:
  html:
    df-print: kable
engine: knitr
page-layout: full
toc: true
---

This notebook demonstrates how to use the Scienceverse platform API from R.
This is a separate Scienceverse deployment with its own `/jobs` API and
credentials. It is not required to run bibr locally or host `bibr serve`.

For direct bibr API access (synchronous, no queue), see [the bibr API notebook](../../notebooks/r_api_demo.qmd).

**Prerequisites:**

- Platform API running (or deployed at `platform.metacheck.app`)
- A platform API key (`sv_...`) — create one via SSO at the platform dashboard
- R packages: `httr2`, `jsonlite`

## Configuration

```{r}
library(httr2)
library(jsonlite)

`%||%` <- function(x, fallback) if (is.null(x)) fallback else x

# Platform API base URL
API_URL <- Sys.getenv("PLATFORM_API_URL", "https://platform.metacheck.app")

# API key — required environment variable
API_KEY <- Sys.getenv("PLATFORM_API_KEY")

# File to process
FILE_PATH <- Sys.getenv("PAPER_PATH", "paper.pdf")

# Polling settings
POLL_INTERVAL <- 2   # seconds between polls
TIMEOUT       <- 600 # max seconds to wait

stopifnot(
  "Set PLATFORM_API_KEY env var" = nchar(API_KEY) > 0,
  "File must exist" = file.exists(FILE_PATH)
)
```

## 1. Health check

Verify the platform API is running and can reach its backends (Postgres, Redis,
bibr API).

```{r}
health <- request(API_URL) |>
  req_url_path_append("health") |>
  req_auth_bearer_token(API_KEY) |>
  req_perform() |>
  resp_body_json()

cat("Health:", health$status, "\n")

ready <- request(API_URL) |>
  req_url_path_append("ready") |>
  req_auth_bearer_token(API_KEY) |>
  req_perform() |>
  resp_body_json()

cat("Ready: ", ready$status, "\n")
cat("Checks:", paste(names(ready$checks), ready$checks, sep = "=", collapse = ", "), "\n")
```

## 2. Submit a job

Upload a file to the platform queue. The response returns immediately with a
`job_id` — the file is processed asynchronously by Arq workers.

```{r}
cat("Submitting", basename(FILE_PATH), "...\n")

submit_resp <- request(API_URL) |>
  req_url_path_append("jobs") |>
  req_auth_bearer_token(API_KEY) |>
  req_body_multipart(file = curl::form_file(FILE_PATH)) |>
  req_timeout(60) |>
  req_perform()

job <- resp_body_json(submit_resp)
job_id <- job$job_id

cat("Job ID:", job_id, "\n")
cat("Status:", job$status, "\n")
```

## 3. Poll for completion

The platform tracks job progress through stages: `reading_upload` →
`processing` → `complete` (or `error:*` on failure).

```{r}
elapsed <- 0

repeat {
  Sys.sleep(POLL_INTERVAL)
  elapsed <- elapsed + POLL_INTERVAL

  status_resp <- request(API_URL) |>
    req_url_path_append("jobs", job_id) |>
    req_auth_bearer_token(API_KEY) |>
    req_timeout(30) |>
    req_perform()

  status <- resp_body_json(status_resp)
  stage <- status$stage %||% ""

  cat(sprintf("[%5.1fs] %s %s\n", elapsed, status$status, stage))

  if (identical(status$status, "complete")) break
  if (identical(status$status, "failed")) stop("Job failed: ", stage)
  if (elapsed >= TIMEOUT) stop("Timed out after ", TIMEOUT, "s")
}
```

## 4. Download the result

Results are returned as structured JSON.

```{r}
result_resp <- request(API_URL) |>
  req_url_path_append("jobs", job_id, "result") |>
  req_url_query(format = "json") |>
  req_auth_bearer_token(API_KEY) |>
  req_timeout(120) |>
  req_perform()

data <- resp_body_json(result_resp)
cat("Downloaded JSON result\n")
```

### Helper: read JSON into named list

```{r}
source("../../scripts/read_json_response.R")
```

### Parse the result

```{r}
result <- parse_bibr_json(resp_body_string(result_resp))
cat("bibr version:", result$info$bibr_version, "\n")
```

## 5. Explore the extracted data

### Paper info

```{r}
result$info
```

### Authors

```{r}
result$author
```

### Sections

```{r}
result$section
```

### Text (sentences)

```{r}
cat("Total sentences:", nrow(result$text), "\n\n")
head(result$text, 10)
```

### References

```{r}
cat("Total references:", nrow(result$bib), "\n\n")
head(result$bib, 10)
```

### URLs

```{r}
result$url
```

### Cross-references

```{r}
cat("Total cross-references:", nrow(result$xref), "\n\n")
if (nrow(result$xref) > 0) head(result$xref, 10)
```

### Figures

```{r}
cat("Total figures:", nrow(result$figure), "\n\n")
if (nrow(result$figure) > 0) result$figure
```

## 6. List your recent jobs

The platform stores job history — you can list previous jobs for your account.

```{r}
jobs_resp <- request(API_URL) |>
  req_url_path_append("jobs") |>
  req_url_query(limit = 10) |>
  req_auth_bearer_token(API_KEY) |>
  req_perform() |>
  resp_body_json()

jobs_df <- do.call(rbind, lapply(jobs_resp, function(j) {
  data.frame(
    job_id = j$job_id,
    filename = j$filename %||% NA,
    status = j$status,
    created = j$created_at %||% NA,
    stringsAsFactors = FALSE
  )
}))
jobs_df
```

## 7. Summary

```{r}
cat("File:           ", basename(FILE_PATH), "\n")
cat("Job ID:         ", job_id, "\n")
cat("Processing time:", elapsed, "seconds\n")
cat("Title:          ", result$info$title, "\n")
cat("DOI:            ", result$info$doi, "\n")
cat("Sentences:      ", nrow(result$text), "\n")
cat("Sections:       ", nrow(result$section), "\n")
cat("Authors:        ", nrow(result$author), "\n")
cat("References:     ", nrow(result$bib), "\n")
cat("URLs:           ", nrow(result$url), "\n")
cat("Cross-refs:     ", nrow(result$xref), "\n")
cat("Figures:        ", nrow(result$figure), "\n")
cat("Tables:         ", nrow(result$table), "\n")
```
